Bird beak's elaboration and exploration

### 2022-10-17 Analysis/03-Elaboration_innovation_analysis.Rmd
Fromaly added a variance test part

## Get the variance for each level
variance_per_clades <- MCMCglmm.variance(shapespace_allbirds_lvl_superorder_order, scale = TRUE)
## Rename the columns
colnames(variance_per_clades) <- c(levels(shapespace$level1)[-1], levels(shapespace$level2)[-1], "phylogeny", "residuals")
reorder_clades <- c(9:35,1:8, 36, 37)
## Pull by term
variance_per_levels<- cbind("orders" = apply(variance_per_clades[, 9:35], 1, sum),
                            "super-orders" = apply(variance_per_clades[, 1:8], 1, sum),
                            "phylogeny" = variance_per_clades[, 36],
                            "residuals" = variance_per_clades[, 37])
variance_per_terms <- cbind("clades" = apply(variance_per_clades[, 1:35], 1, sum),
                            "phylogeny" = variance_per_clades[, 36],
                            "residuals" = variance_per_clades[, 37])
## Plot the results
pdf(file = "../Manuscript/Figures/variance_per_levels_whole_tree.pdf", height = 8, width = 14)
par(mar = c(5,8,2,2), mfrow = c(1, 3))
boxplot(variance_per_terms, horizontal = TRUE, las = 1, main = "Scaled variance by terms", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
boxplot(variance_per_levels, horizontal = TRUE, las = 1, main = "Scaled variance by levels", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
boxplot(variance_per_clades[,reorder_clades], horizontal = TRUE, las = 1, main = "Scaled variance by clades", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
dev.off()

@@

order_shapespace <- MCMCglmm.subsets(n = 4000, # Just run it on 1k matrices for now
      data          = shapespace,
      posteriors    = shapespace_allbirds_lvl_superorder_order,
      group         = MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order)[c(order_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace$level2)[-1], "phylogeny")) #[-1] is because the first level is ""
## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)

## Get the number of observations per groups
variance_per_clades <- MCMCglmm.variance(shapespace_allbirds_lvl_superorder_order, scale = FALSE)
colnames(variance_per_clades) <- c(levels(shapespace$level1)[-1], levels(shapespace$level2)[-1], "phylogeny", "residuals")

## Converting into standard errors
get.se <- function(var, n) {
    sqrt(var[[1]])/sqrt(n)
}
## Get the se per group
se_per_clades <- mapply(get.se, apply(variance_per_clades, 2, list), as.list(c(size.subsets(superorder_shapespace)[-9], size.subsets(order_shapespace), "residuals" = 8748)))
## Scale that
se_per_clades <- se_per_clades/rowSums(se_per_clades)
## Get the other levels
se_per_levels<- cbind("orders" = apply(se_per_clades[, 9:35], 1, sum),
                      "super-orders" = apply(se_per_clades[, 1:8], 1, sum),
                      "phylogeny" = se_per_clades[, 36],
                      "residuals" = se_per_clades[, 37])
se_per_terms <- cbind("clades" = apply(se_per_clades[, 1:35], 1, sum),
                      "phylogeny" = se_per_clades[, 36],
                      "residuals" = se_per_clades[, 37])

## Plot the results
pdf(file = "../Manuscript/Figures/se_per_levels_whole_tree.pdf", height = 8, width = 14)
par(mar = c(5,8,2,2), mfrow = c(1, 3))
boxplot(se_per_terms, horizontal = TRUE, las = 1, main = "Scaled standard error by terms", xlab = "Scaled se ")
boxplot(se_per_levels, horizontal = TRUE, las = 1, main = "Scaled standard error by levels", xlab = "Scaled se")
boxplot(se_per_clades[,reorder_clades], horizontal = TRUE, las = 1, main = "Scaled standard error by clades", xlab = "Scaled variance")
dev.off()
```

>>>>

@@

order_shapespace <- MCMCglmm.subsets(n = 4000, # Just run it on 1k matrices for now
      data          = shapespace,
      posteriors    = shapespace_allbirds_lvl_superorder_order,
      group         = MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order)[c(order_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace$level2)[-1], "phylogeny")) #[-1] is because the first level is ""
## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)
```

### 2022-10-17 Analysis/11-variance_partitioning.Rmd
Fromaly added a variance test part

>>>>

# Analysing the variance per terms

```{r}
## Get the variance for each level
variance_per_clades <- MCMCglmm.variance(shapespace_allbirds_lvl_superorder_order, scale = TRUE)
## Rename the columns
colnames(variance_per_clades) <- c(levels(shapespace$level1)[-1], levels(shapespace$level2)[-1], "phylogeny", "residuals")
reorder_clades <- c(9:35,1:8, 36, 37)
# ## Pull by term
# variance_per_levels<- cbind("orders" = apply(variance_per_clades[, 9:35], 1, sum),
#                             "super-orders" = apply(variance_per_clades[, 1:8], 1, sum),
#                             "phylogeny" = variance_per_clades[, 36],
#                             "residuals" = variance_per_clades[, 37])
# variance_per_terms <- cbind("clades" = apply(variance_per_clades[, 1:35], 1, sum),
#                             "phylogeny" = variance_per_clades[, 36],
#                             "residuals" = variance_per_clades[, 37])
# ## Plot the results
# pdf(file = "../Manuscript/Figures/variance_per_levels_whole_tree.pdf", height = 8, width = 14)
# par(mar = c(5,8,2,2), mfrow = c(1, 3))
# # boxplot(variance_per_terms, horizontal = TRUE, las = 1, main = "Scaled variance by terms", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
# # boxplot(variance_per_levels, horizontal = TRUE, las = 1, main = "Scaled variance by levels", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
# boxplot(variance_per_clades[,reorder_clades], horizontal = TRUE, las = 1, main = "Scaled variance by clades", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
# boxplot(log(variance_per_clades[,reorder_clades])/log(sample_sizes), horizontal = TRUE, las = 1, main = "Scaled variance by clades (removing sample size effect)", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
# boxplot(log(variance_per_clades[,reorder_clades])/log(observed_variance)[names(variances)], horizontal = TRUE, las = 1, main = "Scaled variance by clades (removing variance effect)", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
# dev.off()

# ## Getting the variances per clades
# variances <- apply(variance_per_clades[,reorder_clades], 2, mean)

# ## Getting the sample sizes per clades (the one for the residuals is the whole tree size)
# sample_sizes <- c("residuals" = size.subsets(order_shapespace)[28],
#                   size.subsets(superorder_shapespace),
#                   size.subsets(order_shapespace))[names(variances)]

# ## Measuring the variance for each clade and for the residuals
# ## the obs variance of the residuals is just the variance in the whole dataset (phylogeny)
# observed_variance <- c("residuals" = unlist(get.disparity(dispRity(order_shapespace, metric = sd)))[28], 
#                        unlist(get.disparity(dispRity(superorder_shapespace, metric = sd))),
#                        unlist(get.disparity(dispRity(order_shapespace, metric = sd)))[-28])^2

# plot(log(variances)/log(sample_sizes), log(sample_sizes), xlab = "log variance / log n", ylab = "log n")

## Standard error per terms

```{r}

## Get the number of observations per groups
variance_per_clades <- MCMCglmm.variance(shapespace_allbirds_lvl_superorder_order, scale = FALSE)
colnames(variance_per_clades) <- c(levels(shapespace$level1)[-1], levels(shapespace$level2)[-1], "phylogeny", "residuals")

## Converting into standard errors
get.se <- function(var, n) {
    sqrt(var[[1]])/sqrt(n)
}
## Get the se per group
se_per_clades <- mapply(get.se, apply(variance_per_clades, 2, list), as.list(c(size.subsets(superorder_shapespace)[-9], size.subsets(order_shapespace), "residuals" = 8748)))
## Scale that
se_per_clades <- se_per_clades/rowSums(se_per_clades)
## Get the other levels
se_per_levels<- cbind("orders" = apply(se_per_clades[, 9:35], 1, sum),
                      "super-orders" = apply(se_per_clades[, 1:8], 1, sum),
                      "phylogeny" = se_per_clades[, 36],
                      "residuals" = se_per_clades[, 37])
se_per_terms <- cbind("clades" = apply(se_per_clades[, 1:35], 1, sum),
                      "phylogeny" = se_per_clades[, 36],
                      "residuals" = se_per_clades[, 37])

## Plot the results
pdf(file = "../Manuscript/Figures/se_per_levels_whole_tree.pdf", height = 8, width = 14)
par(mar = c(5,8,2,2), mfrow = c(1, 3))
boxplot(se_per_terms, horizontal = TRUE, las = 1, main = "Scaled standard error by terms", xlab = "Scaled se ")
boxplot(se_per_levels, horizontal = TRUE, las = 1, main = "Scaled standard error by levels", xlab = "Scaled se")
boxplot(se_per_clades[,reorder_clades], horizontal = TRUE, las = 1, main = "Scaled standard error by clades", xlab = "Scaled variance")
dev.off()
```
